Classification structurée pour l'apprentissage par renforcement

### 2012-09-14 klein2012structured_RIA/scirl_ria.tex
Sous section analyse traduite.

\subsection{Analysis}
\label{subsec:scirl:analysis}

In this section, we show that the expert policy $\pi_E$ is close to
be optimal according to the reward function $\R_{\theta_c}$, more
precisely that
$\E_{s\sim\rho_E}[v^*_{\theta_c}(s)-v^{\pi_E}_{\theta_c}(s)]$ is
small. Before stating our main result, we need to introduce some
notations et to define some objects.

We will use the first order discounted future state distribution
concentration coefficient $C_f$~\cite{Munos:2007}:
\begin{equation}
  C_f = (1-\gamma)\sum_{t\geq 0} \gamma^t c(t) \text{ with } c(t) =
  \max_{\pi_1,\dots,\pi_t,s\in\s}\frac{(\rho_E^\top P_{\pi_1}\dots
  P_{\pi_t})(s)}{\rho_E(s)}.
\end{equation}
We note $\pi_c$ the decision rule of the classifier: $\pi_c(s) \in
\argmax_{a\in \A} \theta_c^\top\hat{\mu}^{\pi_E}(s,a)$. The
classification error is therefore $\epsilon_c =
\E_{s\sim\rho_E}[\chi_{\{\pi_c(s)\neq\pi_E(s)\}}] \in [0,1]$. We
write $\hat{Q}^{\pi_E}_{\theta_c} = \theta_c^\top \hat{\mu}^{\pi_E}$
the score function computed from the training set $\D$ (which can be
interpreted as an approximate action-value function). Let also
$\epsilon_{\mu} = \hat{\mu}^{\pi_E} - \mu^{\pi_E}:\s\times\A
\rightarrow  \mathbb{R}^p$ be the feature expectation error.
Consequently, we define the action-value function error as
$\epsilon_Q = \hat{Q}^{\pi_E}_{\theta_c} - Q^{\pi_E}_{\theta_c} =
\theta_c^\top(\hat{\mu}^{\pi_E} - \mu^{\pi_E}) = \theta_c^\top
\epsilon_\mu:\s\times\A\rightarrow\mathbb{R}$. We finally define the
mean delta-max action-value function error as $\bar{\epsilon}_Q =
\E_{s\sim\rho_E}[\max_{a\in\A}\epsilon_Q(s,a) -
\min_{a\in\A}\epsilon_Q(s,a)]\geq 0$.

\begin{theorem}
  \label{th}
  Let $\R_{\theta_c}$ be the reward function outputted by
  Alg.~\ref{algo:scirl}. Let also the quantities $C_f$, $\epsilon_c$
  et $\bar{\epsilon}_Q$ be defined as above. We have
  %
%  the concentration coefficient $C_f$,
%  the classification error $\epsilon_c$ et the mean-max
%  action-value function error $\bar{\epsilon}_Q$ be defined as
%  above. Then, we have
  \begin{equation}
    0\leq
    \E_{s\sim\rho_E}[v^*_{\R_{\theta_c}}-v^{\pi_E}_{\R_{\theta_c}}]
    \leq \frac{C_f}{1-\gamma}\left(\bar{\epsilon}_Q +
    \epsilon_c\frac{2\gamma\|\R_{\theta_c}\|_\infty}{1-\gamma}
%    \epsilon_c\frac{2\gamma}{1-\gamma}\min(\|\R_{\theta_c}\|_\infty,
%    C_p \E_{s\sim\rho_E}[\R_{\theta_c}(s)])
    \right).
  \end{equation}
\end{theorem}
\begin{proof}
  As the proof only relies on the reward $\R_{\theta_c}$, we
  omit the related subscripts to keep the notations simple
  (\textit{e.g.}, $v^\pi$ for
  $v^\pi_{\theta_c}=v^\pi_{\R_{\theta_c}}$ or $\R$ for $\R_{\theta_c}$). First, we link
  the error $\E_{s\sim\rho_E}[v^*(s)-v^{\pi_E}(s)]$ to the
  Bellman residual $\E_{s\sim\rho_E}[[T^*v^{\pi_E}](s)-v^{\pi_E}(s)]$.
  %
  % A second step consists in bounding this residual.
  %
  Componentwise, we have that:
  \begin{align}
    v^* - v^{\pi_E} &= T^* v^*  - T^{\pi^*}v^{\pi_E} +
    T^{\pi^*}v^{\pi_E} - T^* v^{\pi_E} + T^* v^{\pi_E} - v^{\pi_E}
    \\
    &\stackrel{(a)}{\leq} \gamma P_{\pi^*}(v^*-v^{\pi_E}) + T^*
    v^{\pi_E} - v^{\pi_E}
    %
    \stackrel{(b)}{\leq} (I-\gamma
    P_{\pi^*})^{-1} (T^* v^{\pi_E} - v^{\pi_E}).
  \end{align}
  Inequality $(a)$ holds because $T^{\pi^*} v^{\pi_E}\leq T^*
  v^{\pi_E}$ et inequality $(b)$ holds thanks
  to~\cite[Lemma~4.2]{Munos:2007}. Moreover, $v^*$ being optimal we
  have that $v^*-v^{\pi_E}\geq 0$ et $T^*$ being the Bellman
  optimality operator, we have $T^* v^{\pi_E}\geq
  T^{\pi_E}v^{\pi_E}=v^{\pi_E}$. Additionally, remark that
  $(I-\gamma P_{\pi^*})^{-1} = \sum_{t\geq 0}\gamma^t P_{\pi^*}^t$.
  Therefore, using the definition of the concentration coefficient
  $C_f$, we have that:
  \begin{equation}
    0\leq\E_{s\sim\rho_E}[v^*(s)-v^{\pi_E}(s)] \leq \frac{C_f}{1-\gamma}
    \E_{s\sim\rho_E}\left[[T^*v^{\pi_E}](s) - v^{\pi_E}(s)\right].
    \label{eq:proof:residual}
  \end{equation}
  This results actually follows closely the one
  of~\cite[Theorem~4.2]{Munos:2007}. There remains to bound the
  Bellman residual $\E_{s\sim\rho_E}[[T^*v^{\pi_E}](s) -
  v^{\pi_E}(s)]$. Considering the following decomposition,
  \begin{equation}
    T^* v^{\pi_E} - v^{\pi_E} = T^* v^{\pi_E} - T^{\pi_c}v^{\pi_E}
    + T^{\pi_c}v^{\pi_E}- v^{\pi_E},
    \label{eq:proof:decomposition}
  \end{equation}
  we will bound $\E_{s\sim\rho_E}[[T^* v^{\pi_E}](s) - [T^{\pi_c}v^{\pi_E}](s)]$
  et $\E_{s\sim\rho_E}[[T^{\pi_c}v^{\pi_E}](s) - v^{\pi_E}(s)]$.

  The policy $\pi_c$ (the decision rule of the classifier) is greedy
  with respect to
  $\hat{Q}^{\pi_E}=\theta_c^\top\hat{\mu}^{\pi_E}$. Therefore, for any state-action couple
  $(s,a)\in\s\times \A$ we have:
  \begin{equation}
    \hat{Q}^{\pi_E}(s,\pi_c(s))\geq
    \hat{Q}^{\pi_E}(s,a)
    \Leftrightarrow
    Q^{\pi_E}(s,a) \leq Q^{\pi_E}(s,\pi_c(s)) +
    \epsilon_Q(s,\pi_c(s)) - \epsilon_Q(s,a).
  \end{equation}
  By definition, $Q^{\pi_E}(s,a) = [T^a v^{\pi_E}](s)$ and
  $Q^{\pi_E}(s,\pi_c(s)) = [T^{\pi_c} v^{\pi_E}](s)$. Therefore, for $s\in\s$:
  \begin{align}
    \forall a\in A,\; [T^a v^{\pi_E}](s) &\leq [T^{\pi_c}
    v^{\pi_E}](s) + \epsilon_Q(s,\pi_c(s))-\epsilon_Q(s,a)
    \\
    \Rightarrow [T^* v^{\pi_E}](s) &\leq [T^{\pi_c}
    v^{\pi_E}](s) + \max_{a\in \A}\epsilon_Q(s,a)-\min_{a\in
    \A}\epsilon_Q(s,a).
  \end{align}
  Taking the expectation according to $\rho_E$ et noticing that
  $T^* v^{\pi_E}\geq v^{\pi_E}$, we bound the first
  term:
  \begin{equation}
    0 \leq \E_{s\sim\rho_E}\left[ [T^* v^{\pi_E}](s) - [T^{\pi_c}
    v^{\pi_E}](s)\right] \leq \bar{\epsilon}_Q.
    \label{eq:proof:b1}
  \end{equation}
  There finally remains to bound the term $\E_{s\sim\rho_E}[[T^{\pi_c}v^{\pi_E}](s) -
  v^{\pi_E}(s)]$.

  Let us write $M\in\mathbb{R}^{|\s|\times |\s|}$ the diagonal matrix
  defined as $M = \diag (\chi_{\{\pi_c(s)\neq\pi_E(s)\}})$. Using
  this, the Bellman operator $T^{\pi_c}$ may be written as, for any $v\in\mathbb{R}^\s$:
  \begin{equation}
    T^{\pi_c}v = \R + \gamma M P_{\pi_c} v + \gamma (I-M)P_{\pi_E} v
    = \R + \gamma P_{\pi_E} v + \gamma M (P_{\pi_c}-P_{\pi_E})v.
  \end{equation}
  Applying this operator to $v^{\pi_E}$ et recalling that $\R +
  \gamma P_{\pi_E} v^{\pi_E} = T^{\pi_E} v^{\pi_E} = v^{\pi_E}$, we
  get:
  \begin{equation}
    T^{\pi_c}v^{\pi_E} - v^{\pi_E} = \gamma M
    (P_{\pi_c}-P_{\pi_E})v^{\pi_E}
    \Rightarrow |\rho_E^\top (T^{\pi_c}v^{\pi_E} - v^{\pi_E})| = \gamma
    |\rho_E^\top M (P_{\pi_c}-P_{\pi_E})v^{\pi_E}|.
  \end{equation}
  One can easily see that $\|(P_{\pi_c}-P_{\pi_E})v^{\pi_E}\|_\infty
  \leq \frac{2}{1-\gamma}\|\R\|_\infty$, which allows bounding the
  last term:
  \begin{equation}
    |\E_{s\sim\rho_E}[[T^{\pi_c}v^{\pi_E}](s) - v^{\pi_E}(s)]| \leq
    \epsilon_c \frac{2\gamma}{1-\gamma} \|\R\|_\infty.
    \label{eq:proof:b2}
  \end{equation}
  Injecting bounds of Eqs.~\eqref{eq:proof:b1}
  and~\eqref{eq:proof:b2} into Eq.~\eqref{eq:proof:residual} gives the
  stated result.%, which concludes the proof.
%  Injecting bounds of Eqs.~\eqref{eq:proof:b1}
%  and~\eqref{eq:proof:b2} into Eq.~\eqref{eq:proof:residual} using the
%  decomposition of Eq.~\eqref{eq:proof:decomposition} gives the
%  stated result:
%  \begin{equation}
%    0\leq\E_{s\sim\rho_E}[v^*(s)-v^{\pi_E}(s)] \leq
%    \frac{C_f}{1-\gamma} \left(\bar{\epsilon}_Q + \epsilon_c \frac{2\gamma}{1-\gamma} \|\R\|_\infty
%    \right).
%  \end{equation}
%  This concludes the proof.
\end{proof}

This result shows that if the expert feature expectation is well
estimated (in the sense that the estimation error $\epsilon_\mu$ is
small for states sampled according to the expert stationary policy
and for all actions) et if the classification error $\epsilon_c$ is
small, then the proposed generic algorithm outputs a reward function
$\R_{\theta_c}$ for which the expert policy will be near optimal. A
direct corollary of Th.~\ref{th} is that given the true expert
feature expectation $\mu^{\pi_E}$ et a perfect classifier
($\epsilon_c=0$), $\pi_E$ is the unique optimal policy for
$\R_{\theta_c}$.

One may argue that this bounds trivially holds for the null reward
function (a reward often exhibited to show that IRL is an ill-posed
problem), obtained if $\theta_c=0$. However, recall that the
parameter vector $\theta_c$ is computed by the classifier. With
$\theta_c=0$, the decision rule would be a random policy et we
would have $\epsilon_c = \frac{|\A|-1}{|\A|}$, the worst possible
classification error. This case is really unlikely. Therefore, we
advocate that the proposed approach somehow allows disambiguating
the IRL problem (at least, it does not output trivial reward
functions such as the null vector).
%
Also, this bound is scale-invariant: one could impose
$\|\theta_c\|=1$ or normalize (action-) value functions by
$\|\R_{\theta_c}\|_\infty^{-1}$.

One should notice that there is a hidden dependency of the
classification error $\epsilon_c$ to the estimated expert feature
expectation $\hat{\mu}^{\pi_E}$. Indeed, the minimum classification
error depends on the hypothesis space spanned by the chosen score
function basis functions for the MC$^2$ algorithm (here
$\hat{\mu}^{\pi_E}$). Nevertheless, provided a good representation
for the reward function (that is a good choice of basis functions
$\phi_i$) et a small estimation error, this should not be a
practical problem.

Finally, if our bound relies on the generalization errors
$\epsilon_c$ et $\bar{\epsilon}_Q$, the classifier will only use
$(\hat{\mu}^{\pi_E}(s_i,a))_{1\leq i\leq N,a\in\A}$ in the training
phase, where $s_i$ are the states from the set $\D$. It outputs
$\theta_c$, seen as a reward function, thus the estimated feature
expectation $\hat{\mu}^{\pi_E}$ is no longer required. Therefore,
practically it should be sufficient to estimate well
$\hat{\mu}^{\pi_E}$ on state-action couples $(s_i,a)_{1\leq i\leq
N,a\in\A}$, which allows envisioning Monte-Carlo rollouts for
example.

>>>>

\subsection{Analyse}
\label{subsec:scirl:analysis}

Dans cette section, nous montrons que la politique de l'expert $\pi_E$ est quasi-optimale vis-à-vis de la fonction de récompense $\R_{\theta_c}$, plus précisément que 
$\E_{s\sim\rho_E}[v^*_{\theta_c}(s)-v^{\pi_E}_{\theta_c}(s)]$ est faible.
Avant de présenter le résultat principal, il nous faut introduire quelques notations et définir quelques objets.

Nous allons utiliser le coefficient de concentration de la distribution pondérée du premier ordre pour l'état futur $C_f$~\cite{Munos:2007}:
\begin{equation}
  C_f = (1-\gamma)\sum_{t\geq 0} \gamma^t c(t) \text{ with } c(t) =
  \max_{\pi_1,\dots,\pi_t,s\in\s}\frac{(\rho_E^\top P_{\pi_1}\dots
  P_{\pi_t})(s)}{\rho_E(s)}.
\end{equation}
On note $\pi_c$ la règle de décision du classifieur : $\pi_c(s) \in
\argmax_{a\in \A} \theta_c^\top\hat{\mu}^{\pi_E}(s,a)$. L'erreur de classification est donc $\epsilon_c =
\E_{s\sim\rho_E}[\chi_{\{\pi_c(s)\neq\pi_E(s)\}}] \in [0,1]$. On écrit $\hat{Q}^{\pi_E}_{\theta_c} = \theta_c^\top \hat{\mu}^{\pi_E}$ la fonction de score calculée à partir du set d'entraînement $\D$ (celle-ci peut être interprétée comme une fonction de qualité approchée). Soit 
$\epsilon_{\mu} = \hat{\mu}^{\pi_E} - \mu^{\pi_E}:\s\times\A
\rightarrow  \mathbb{R}^p$ l'erreur sur l'estimation de l'attribut moyen.
En conséquence, on définit l'erreur sur la fonction de qualité par 
$\epsilon_Q = \hat{Q}^{\pi_E}_{\theta_c} - Q^{\pi_E}_{\theta_c} =
\theta_c^\top(\hat{\mu}^{\pi_E} - \mu^{\pi_E}) = \theta_c^\top
\epsilon_\mu:\s\times\A\rightarrow\mathbb{R}$. Finalement, on définit le delta-max moyen de l'erreur sur la fonction de qualité par $\bar{\epsilon}_Q =
\E_{s\sim\rho_E}[\max_{a\in\A}\epsilon_Q(s,a) -
\min_{a\in\A}\epsilon_Q(s,a)]\geq 0$.

\begin{theorem}
  \label{th}
  Soit $\R_{\theta_c}$ la fonction de récompense renvoyée par l'Alg.~\ref{algo:scirl}. Soient $C_f$, $\epsilon_c$
  et $\bar{\epsilon}_Q$ les quantités définies ci-dessus. On a
  %
%  the concentration coefficient $C_f$,
%  the classification error $\epsilon_c$ et the mean-max
%  action-value function error $\bar{\epsilon}_Q$ be defined as
%  above. Then, we have
  \begin{equation}
    0\leq
    \E_{s\sim\rho_E}[v^*_{\R_{\theta_c}}-v^{\pi_E}_{\R_{\theta_c}}]
    \leq \frac{C_f}{1-\gamma}\left(\bar{\epsilon}_Q +
    \epsilon_c\frac{2\gamma\|\R_{\theta_c}\|_\infty}{1-\gamma}
%    \epsilon_c\frac{2\gamma}{1-\gamma}\min(\|\R_{\theta_c}\|_\infty,
%    C_p \E_{s\sim\rho_E}[\R_{\theta_c}(s)])
    \right).
  \end{equation}
\end{theorem}
\begin{proof}
  La démonstration repose uniquement sur la récompense $\R_{\theta_c}$, on omet donc par soucis de clarté des notations certains indices relatifs à la récompense 
  (\textit{e.g.}, $v^\pi$ pour
  $v^\pi_{\theta_c}=v^\pi_{\R_{\theta_c}}$ ou $\R$ pour $\R_{\theta_c}$). Tout d'abord, l'on relie  l'erreur $\E_{s\sim\rho_E}[v^*(s)-v^{\pi_E}(s)]$ au résidu de Bellman $\E_{s\sim\rho_E}[[T^*v^{\pi_E}](s)-v^{\pi_E}(s)]$.
  %
  % A second step consists in bounding this residual.
  %
  Composante par composante, on a :
  \begin{align}
    v^* - v^{\pi_E} &= T^* v^*  - T^{\pi^*}v^{\pi_E} +
    T^{\pi^*}v^{\pi_E} - T^* v^{\pi_E} + T^* v^{\pi_E} - v^{\pi_E}
    \\
    &\stackrel{(a)}{\leq} \gamma P_{\pi^*}(v^*-v^{\pi_E}) + T^*
    v^{\pi_E} - v^{\pi_E}
    %
    \stackrel{(b)}{\leq} (I-\gamma
    P_{\pi^*})^{-1} (T^* v^{\pi_E} - v^{\pi_E}).
  \end{align}
  L'inégalité $(a)$ est valable car $T^{\pi^*} v^{\pi_E}\leq T^*
  v^{\pi_E}$ et l'inégalité $(b)$ l'est aussi grâce à~\cite[Lemma~4.2]{Munos:2007}. De plus, $v^*$ étant optimale, l'on voit que  $v^*-v^{\pi_E}\geq 0$ et avec $T^*$ l'opérateur d'optimalité de Bellman, on a $T^* v^{\pi_E}\geq
  T^{\pi_E}v^{\pi_E}=v^{\pi_E}$. De plus, on remarque que 
  $(I-\gamma P_{\pi^*})^{-1} = \sum_{t\geq 0}\gamma^t P_{\pi^*}^t$.
  Donc, d'après la définition du coefficient de concentration 
  $C_f$, on a :
  \begin{equation}
    0\leq\E_{s\sim\rho_E}[v^*(s)-v^{\pi_E}(s)] \leq \frac{C_f}{1-\gamma}
    \E_{s\sim\rho_E}\left[[T^*v^{\pi_E}](s) - v^{\pi_E}(s)\right].
    \label{eq:proof:residual}
  \end{equation}
  Ce résultat est similaire à celui de~\cite[Theorem~4.2]{Munos:2007}. Il reste à borner le résidu de Bellman $\E_{s\sim\rho_E}[[T^*v^{\pi_E}](s) -
  v^{\pi_E}(s)]$. Considérons la décomposition
  \begin{equation}
    T^* v^{\pi_E} - v^{\pi_E} = T^* v^{\pi_E} - T^{\pi_c}v^{\pi_E}
    + T^{\pi_c}v^{\pi_E}- v^{\pi_E},
    \label{eq:proof:decomposition}
  \end{equation}
  on va borner  $\E_{s\sim\rho_E}[[T^* v^{\pi_E}](s) - [T^{\pi_c}v^{\pi_E}](s)]$
  et $\E_{s\sim\rho_E}[[T^{\pi_c}v^{\pi_E}](s) - v^{\pi_E}(s)]$.

  La politique $\pi_c$ (la règle de décision du classifieur) est gloutonne vis-à-vis de 
  $\hat{Q}^{\pi_E}=\theta_c^\top\hat{\mu}^{\pi_E}$. Donc, pour chaque couple état-action
  $(s,a)\in\s\times \A$ on a:
  \begin{equation}
    \hat{Q}^{\pi_E}(s,\pi_c(s))\geq
    \hat{Q}^{\pi_E}(s,a)
    \Leftrightarrow
    Q^{\pi_E}(s,a) \leq Q^{\pi_E}(s,\pi_c(s)) +
    \epsilon_Q(s,\pi_c(s)) - \epsilon_Q(s,a).
  \end{equation}
  Par définition, $Q^{\pi_E}(s,a) = [T^a v^{\pi_E}](s)$ et
  $Q^{\pi_E}(s,\pi_c(s)) = [T^{\pi_c} v^{\pi_E}](s)$. Donc, pour $s\in\s$:
  \begin{align}
    \forall a\in A,\; [T^a v^{\pi_E}](s) &\leq [T^{\pi_c}
    v^{\pi_E}](s) + \epsilon_Q(s,\pi_c(s))-\epsilon_Q(s,a)
    \\
    \Rightarrow [T^* v^{\pi_E}](s) &\leq [T^{\pi_c}
    v^{\pi_E}](s) + \max_{a\in \A}\epsilon_Q(s,a)-\min_{a\in
    \A}\epsilon_Q(s,a).
  \end{align}
  En passant à l'espérance selon $\rho_E$ et tout en remarquant que 
  $T^* v^{\pi_E}\geq v^{\pi_E}$, on borne le premier terme
  \begin{equation}
    0 \leq \E_{s\sim\rho_E}\left[ [T^* v^{\pi_E}](s) - [T^{\pi_c}
    v^{\pi_E}](s)\right] \leq \bar{\epsilon}_Q.
    \label{eq:proof:b1}
  \end{equation}
  Il reste enfin à borner le terme $\E_{s\sim\rho_E}[[T^{\pi_c}v^{\pi_E}](s) -
  v^{\pi_E}(s)]$.

  On écrit $M\in\mathbb{R}^{|\s|\times |\s|}$ la matrice diagonale définie par $M = \diag (\chi_{\{\pi_c(s)\neq\pi_E(s)\}})$. En utilisant cela, l'opérateur de Bellman $T^{\pi_c}$ peut s'écrire, pour tout $v\in\mathbb{R}^\s$:
  \begin{equation}
    T^{\pi_c}v = \R + \gamma M P_{\pi_c} v + \gamma (I-M)P_{\pi_E} v
    = \R + \gamma P_{\pi_E} v + \gamma M (P_{\pi_c}-P_{\pi_E})v.
  \end{equation}
  En appliquant cet opérateur à  $v^{\pi_E}$ et en utilisant le fait que $\R +
  \gamma P_{\pi_E} v^{\pi_E} = T^{\pi_E} v^{\pi_E} = v^{\pi_E}$, on obtient :
  \begin{equation}
    T^{\pi_c}v^{\pi_E} - v^{\pi_E} = \gamma M
    (P_{\pi_c}-P_{\pi_E})v^{\pi_E}
    \Rightarrow |\rho_E^\top (T^{\pi_c}v^{\pi_E} - v^{\pi_E})| = \gamma
    |\rho_E^\top M (P_{\pi_c}-P_{\pi_E})v^{\pi_E}|.
  \end{equation}
  On peut facilement voir que $\|(P_{\pi_c}-P_{\pi_E})v^{\pi_E}\|_\infty
  \leq \frac{2}{1-\gamma}\|\R\|_\infty$, ce qui permet de borner le dernier terme
  \begin{equation}
    |\E_{s\sim\rho_E}[[T^{\pi_c}v^{\pi_E}](s) - v^{\pi_E}(s)]| \leq
    \epsilon_c \frac{2\gamma}{1-\gamma} \|\R\|_\infty.
    \label{eq:proof:b2}
  \end{equation}
  Injecter les bornes des Eqs.~\eqref{eq:proof:b1}
  and~\eqref{eq:proof:b2} dans l'Eq.~\eqref{eq:proof:residual} montre le résultat proposé.%, which concludes the proof.
%  Injecting bounds of Eqs.~\eqref{eq:proof:b1}
%  and~\eqref{eq:proof:b2} into Eq.~\eqref{eq:proof:residual} using the
%  decomposition of Eq.~\eqref{eq:proof:decomposition} gives the
%  stated result:
%  \begin{equation}
%    0\leq\E_{s\sim\rho_E}[v^*(s)-v^{\pi_E}(s)] \leq
%    \frac{C_f}{1-\gamma} \left(\bar{\epsilon}_Q + \epsilon_c \frac{2\gamma}{1-\gamma} \|\R\|_\infty
%    \right).
%  \end{equation}
%  This concludes the proof.
\end{proof}

Ce résultat montre que si l'attribut moyen de l'expert est bien estimé (dans le sens d'une faible erreur d'estumation $\epsilon_\mu$ pour les états échantillonnés selon la politique stationnaire de l'expert et pour toutes les actions) et si l'erreur de classification $\epsilon_c$ et également faible alors l'algorithme générique proposé fournit une fonction de récompense 
$\R_{\theta_c}$ vis-à-vis de laquelle l'expert sera quasi-optimal. Un corollaire direct du Th.~\ref{th} stipule qu'avec le vrai attribut moyen de l'expert $\mu^{\pi_E}$ et un classifieur parfait 
($\epsilon_c=0$), $\pi_E$ est l'unique politique optimale pour 
$\R_{\theta_c}$.

L'on pourrait arguer que ces bornes sont triviallement valables pour la fonction de récompense nulle (fonction régulièrement citée en exemple de la nature mal posée du problème de l'ARI) correspondant au cas $\theta_c=0$. Cependant il faut se rappeler que le vecteur de paramètres $\theta_c$ est choisi par le classifieur. Avec 
$\theta_c=0$, la règle de décision serait une politique aléatoire et nous aurions $\epsilon_c = \frac{|\A|-1}{|\A|}$, c'est-à-dire la pire erreur de classification possible. Ce cas est donc très improbable. De fait nous affirmons que notre approche permet, d'une certaine manière, de lever l'ambiguité du problème de l'ARI (au moins, l'algorithme ne renvoie pas de récompense triviales comme la récompense nulle).
%
On peut également noter que cette borne est invariante par dilatation. L'on pourrait imposer 
$\|\theta_c\|=1$ ou normaliser la fonction de valeur (ou de qualité) par 
$\|\R_{\theta_c}\|_\infty^{-1}$.

On remarquera une dépendance caché de l'erreur de classification  $\epsilon_c$ à l'estimation de l'attribut moyen de l'expert  $\hat{\mu}^{\pi_E}$. En effet, l'erreur de classification minimale dépend de l'espace d'hypothèses généré par les fonctions de base de la fonction de score de l'algorithme de MC$^2$ (ici, 
$\hat{\mu}^{\pi_E}$). Néanmoins, avec une bonne représentation de la fonction de récompense (c'est-à-dire un choix judicieux de fonctions de base $\phi_i$) et une faible erreur d'estimation, cela ne devrait poser aucun problème en pratique.

Finalement, si notre borne se base sur les erreurs en généralisation 
$\epsilon_c$ et $\bar{\epsilon}_Q$, le classifieur n'utilisera 
$(\hat{\mu}^{\pi_E}(s_i,a))_{1\leq i\leq N,a\in\A}$ que lors de la phase d'entraînement, où les $s_i$ sont les états présents dans $\D$. Il renvoie 
$\theta_c$, vu comme une fonction de récompense, donc l'estimée de l'attribut moyen $\hat{\mu}^{\pi_E}$ n'est plus nécessaire. Donc en pratique il est satisfaisant d'estimer 
$\hat{\mu}^{\pi_E}$ correctement pour les couples état-action $(s_i,a)_{1\leq i\leq
N,a\in\A}$, ce qui permet d'envisager une simple estimation de Monte-Carlo par exemple.
